---
title: "Kat_Waldo_ESM206_Ass4_Lobsters"
author: "Kat Leigh"
date: "11/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

INTRODUCTION:
This report analyzes differences in the abundance and size of California Spiny Lobsters (Panulirus interruptus) observed at five locations along the Santa Barbara coastline by the Santa Barbara Coastal Long Term Ecological Research program (SBC LTER). Spiny lobsters play a critical predatory role in Southern California's giant kelp forests, but are threatened by fishing pressure. The California Fish and Game's Network of Marine Protected Areas (MPAs), Naples and Isla Vista, provide a refuge from fishing pressure and thus a unique oportunity to compare and contrast lobsters living in communities subject to and sheltered from fishing. To this end, this report examines:
- A: changes in annual lobster abundance (counts) over time by site;
- B: shifts in lobster size distributions between 2012 and 2018; and
- C: differences between mean lobster sizes at MPA vs. non-MPA sites in 2012 and 2018.

As a result of this examination, it was found that:
- between MPA and non-MPA sites, there was XXX significant difference in lobster size in 2012, and xxx significant difference in lobster size in 2018;
- within MPA sites, there was xxx significant mean size difference between lobsters observed in 2012 and 2018; and
- within non-MPA sites, there was xxx significant mean size difference in lobsters observed between 2012 and 2018.

These findings provide insight into kelp forest management, indicating that MPAs could be xxx effective at helping support healthy lobster populations.

DATA AND METHODS: 
The data used in this report, from the SBC LTER's two study reefs located in or near the Naples and Isla Vista MPAs, ranges from 2012 to present (2019). Observations were collected annually by divers in late summer before the start of the fishing season from two sites located within the MPAs and three located outside of the MPAs (Arroyo Quemado, Mohawk and Carpinteria). The dependent variables examined were time (the year of observation), and locaton (the site), and the independent variables examined were size (the carapace length in millimeters ) and count (the number of lobsters). Statistical difference was assessed using a two-sided t-test with a Welsh assumption, and a level of significance of 0.05. Analysis was performed using Rstudio software version 1.2.1335.

```{r, message = FALSE, warning = FALSE}
# ----
# run libraries and read data
# ----

library(tidyverse)
library(janitor)
library(lubridate)

lobster_df <- read_csv(here::here("data", "lobster_abundance_sbc_lter.csv"), 
                       na = "-99999") %>% 
  clean_names()

tidy_lob <- lobster_df %>% 
  uncount(count)

```

### TITLE

## Results A:

Visually explore changes in annual lobster abundance (counts) by site. After grouping observations to find lobster counts for each year, create a finalized data visualization (no modeling/stats needed) showing changes in annual lobster abundance at the five sites over time.

```{r}
# ----
# Data wrangling
# ----

lob_site_yr <- tidy_lob %>%
  group_by(year) %>% 
  count(site) %>% 
  dplyr::mutate(
    yr = dplyr::case_when(
      year %in% "2012" ~ "'12",
      year %in% "2013" ~ "'13",
      year %in% "2014" ~ "'14",
      year %in% "2015" ~ "'15",
      year %in% "2016" ~ "'16",
      year %in% "2017" ~ "'17",
      year %in% "2018" ~ "'18",
    )
  )
   

# ----
# Data plotting
# ----
  
ggplot(data = lob_site_yr, aes(x = yr, y = n)) +
  geom_col(aes(fill = site), show.legend = FALSE, 
           color = "black",  alpha = 0.8) +
  scale_fill_brewer(palette = "Pastel1") + # Chose this color palette to have color stand out
  scale_y_discrete(expand = c(0,0)) + # Removed bottom space
  labs(x = "Year", y = "Number of Lobsters", title = "Number of Lobsters at Each Site from 2012-2018", fill = "Site") +
  theme_minimal() +
  facet_wrap(~site, scales = "free_x") +
  theme(panel.spacing = unit(2, "lines"))
```
Fig. 1: Lobster Abundance per site for years 2012-2018. Note that both sites Naples (NAPL) and Isla Vista (IVEE) are MPAs.


## Results B:

Visually explore lobster size distribution shifts by comparing lobster sizes in 2012 and 2018. Visually explore the size distributions for lobsters observed at the 5 different sites in 2012, and in 2018 (i.e. you will have a total of 10 size distributions shown: for each of the 5 sites, one for 2012 and one for 2018). Make a finalized graph that clearly and professionally shows community size distributions at each site in the 2 years, allowing the reader to easily interpret distribution shifts.
```{r}
# ----
# Data wrangling
# ----

lob_size_yr <- tidy_lob %>% 
  select(year, site, size_mm) %>% 
  filter(year == c(2012,2018))

# ----
# Data plotting
# ----

ggplot(lob_size_yr) +
  # Made geom_violin with areas proportional to count
  geom_violin(show.legend = FALSE, scale = "count",
              aes(x = site, y = size_mm, fill = site)) + 
  scale_fill_brewer(palette = "Pastel1") +
  facet_wrap(~year) +
  # Removed bottom space and showed full scale
  scale_y_continuous(expand = c(0,0), limits = c(0,200)) + 
  labs(y = "Size [mm]", x = "Site", caption = "Area is proportional to count",
       title = "Lobster size distribution") +
  theme_minimal()
```

## Results C:

Compare mean lobster sizes at MPA vs. non-MPA sites in 2012 and 2018. Here, for each year (2012 and 2018) consolidate the size observations into only two groups: MPA and non-MPA lobsters. Then answer the following four questions: 

For 2012 observations, is there a significant difference in lobster size between MPA and non-MPA sites? 
For 2018 observations, is there a significant difference in lobster size between MPA and non-MPA sites? 
For MPA sites only, is there a significant difference in lobsters observed in 2012 vs. 2018?
For non-MPA sites only, is there a significant difference in lobsters observed in 2012 vs. 2018?

```{r}
# ----
# Data wrangling
# ----

lob_mpa <- lob_size_yr %>% 
  filter(site %in% c("IVEE", "NAPL")) 
  
lob_nmpa <- lob_size_yr %>% 
  filter(site %in% c("AQUE", "CARP", "MOHK")) 

# MPA 2012
sample_mpa_12 <- lob_mpa %>%
  filter(year == 2012) %>% 
  pull(size_mm)

# Non-MPA 2012  
sample_nmpa_12 <- lob_nmpa %>%
  filter(year == 2012) %>% 
  pull(size_mm)

# MPA 2018
sample_mpa_18 <- lob_mpa %>% 
  filter(year == 2018) %>% 
  pull(size_mm)

# Non-MPA 2018
sample_nmpa_18 <- lob_nmpa %>% 
  filter(year == 2018) %>% 
  pull(size_mm)


# 2012 observations mpa vs non-mpa   
mn12_ttest <- t.test(sample_mpa_12, sample_nmpa_12)
mn12_ttest

# 2018 observations mpa vs non-mpa
mn18_ttest <- t.test(sample_mpa_18, sample_nmpa_18)
mn18_ttest

# MPA sites only
mpa_ttest <- t.test(sample_mpa_12, sample_mpa_18)
mpa_ttest

# Non MPA sites only
nmpa_ttest <- t.test(sample_nmpa_12, sample_nmpa_18)
nmpa_ttest

```

